---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tsibble)
library(feasts)
library(plotly)
library(slider)
library(shiny)
theme_set(theme_bw())
color_pal <- c("#B0D1D9", "#273859")
```

[Global Carbon Footprint Network II] This problem creates interactive versions
of the carbon emissions dataset introduced in the previous problem.

a. Using `ggplotly`, create a line plot where country names can be revelaed
on hover.

This code is copied from the previous demo.

```{r}
carbon <- read_csv("https://go.wisc.edu/7qx7u1")
carbon_ts <- carbon |>
  pivot_longer(-c("country", "Country Code"), names_to = "year") |>
  filter(country != "Switzerland") |>
  mutate(year = as.integer(year)) |>
  as_tsibble(index = year, key = country)

feat_acf(carbon_ts$value[1:100], lag_max = 10)

acfs <- list()
countries <- unique(carbon_ts$country)
for (i in seq_along(countries)) {
    acfs[[i]] <- carbon_ts |>
        filter(country == countries[i]) |>
        ACF(value)
}

x <- bind_rows(acfs)
x

smoothed_carbon <- carbon_ts |>
  group_by(country) |>
  mutate(smoothed = slide_dbl(value, ~ mean(.), .before = 5, .after = 5)) |>
  ungroup()

ggplot(smoothed_carbon) +
  geom_line(aes(year, smoothed, group = country))
```

b. Propose an interactive time series visualization that graphically queries a
time series feature (or several features) that can be generated using the
`feasts` package. What questions is your visualization particularly well-suited
to? What questions is it not appropriate for?

```{r}
p <- ggplot(smoothed_carbon) +
  geom_line(aes(year, smoothed, group = country))
ggplotly(p)
p
```

c. Implement your proposal from part (b).

```{r}
ui <- fluidPage(
  fluidRow(
    column(6, plotOutput("scatter", brush = "brush")),
    column(6, plotlyOutput("trends"))
  )
)

server <- function(input, output) {
  selected <- reactiveVal(rep(FALSE, nrow(carbon_trends)))
  observeEvent(
    input$brush, {
        selected(brushedPoints(carbon_trends, input$brush, allRows = TRUE)$selected_)
  })

  output$scatter <- renderPlot({
      carbon_trends |>
          group_by(country) |>
          filter(year == 1961) |>
          ggplot() +
            geom_point(aes(log_spikiness, curvature))
  })

  output$trends <- renderPlotly({
      #browser()
      if (!any(selected())) {
          return (ggplotly(ggplot()))
      }
      p <- carbon_trends |>
          filter(selected()) |>
          ggplot() +
          geom_line(aes(year, value, group = country))
      ggplotly(p)
  })
}

shinyApp(ui, server)
```