---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
th <- theme_classic() +
    theme(
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        strip.text = element_text(size = 20)
    )
theme_set(th)
```


[Plant Growth Experiment] This problem will give you practice with tidying a
dataset so that it can be easily visualized. The data describe the height of
several plants measured every 7 days. The plants have been treated with
different amounts of a growth stimulant. The first few rows are printed below --
`height.x` denotes the height of the plant on day `x`.

```{r}
library(tidyverse)
plants <- read_csv("https://uwmadison.box.com/shared/static/qg9gwk2ldjdtcmmmiropcunf34ddonya.csv")
plants
```

a. Propose an alternative arrangement of rows and columns that conforms to the
tidy data principle.

b. Implement your proposed arrangement from part (a).

```{r}
plants_longer <- plants |>
    pivot_longer(starts_with("height"), names_to = "time", values_to = "height") |>
    mutate(time = as.integer(str_extract(time, "[0-9]+")))

plants_longer

ggplot(plants_longer) +
    geom_line(aes(time, height, col = treatment, group = interaction(treatment, plantid)))

# graphical mark
ggplot(plants_longer) +
    geom_line(aes(time, height, col = treatment, group = interaction(treatment, plantid))) +
    facet_wrap(~ treatment)

# labels? more meaning? X
# when were the measurements?
# title X
# dashes?
# too much overlap -> spread across panels?

ggplot(plants_longer, aes(time, height, col = treatment, )) +
    geom_line(aes(group = interaction(treatment, plantid))) +
    geom_point(size = 5) +
    scale_x_continuous(breaks = c(0, 7, 14, 21)) +
    labs(x = "Time (Days)", y = "Height (cm)", col = "Treatment", title = "Effect of Treatments on Plant Growth") +
    scale_color_manual(values = c("#f15a5a", "#a8dce8", "#5ca0b0", "#145563")) +
    theme(title = element_text(size = 18))

plants_longer |>
    arrange(plantid) |>
    print(n = 20)

# encoding
aes()
x -> time
y -> height
color -> treatment

```


c. Using the dataset from (b), design and implement a visualization showing the
growth of the plants over time according to different treatments.
