---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(glue)
library(tidymodels)
th <- theme_minimal() +
    theme(
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "#f7f7f7"),
        panel.border = element_rect(fill = NA, color = "#0c0c0c"),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        strip.text = element_text(size = 16),
        legend.position = "bottom"
    )
theme_set(th)
```

[Olympics PCA - Continued] Let's study Olympics dataset from the previous
problem. We'll apply PCA ourselves, compare alternative visual encodings, and
clarify the impact of standardization. The code below loads the package.

```{r}
#install.packages("ade4")
data(olympic, package = "ade4")
head(olympic$tab)
```

a. Define a tidymodels recipe that normalizes all features and specifies that
PCA should be performed.

b. Plot and interpret the variance explained by each principal component.

c. Visualize the scores associated with each athlete.

d. Visualize the top 4 principal components.  Pick an athlete from part (c) and
comment on their performance relative to the average decathlete for that year.

e. Rerun the previous steps but with normalization replaced with only scaling
(not centering). How do the results change? Which approach would you recommend
and why?